---
title: "mrgsolve evaluation"
author: "Justin Penzenstadler"
date: "March 3, 2016"
output: html_document
---

##Set up 
There is a protocol to install this package that must be followed exactly. Windows installations of R are not configured to install mrgsolve out of the box as source code.  mrgsolve AND key dependencies (rcpp and RcppArmadillo) must be downloaded as source and compiled on the computer, rather than a typical installatin binary files.  This convention is popular for linuxboxes with many different hardware configs.

##Protocol 
**took JAP 3 hours to debug** derp
- Download latest version of Rtools and run the exe https://cran.r-project.org/bin/windows/Rtools/
  - - YOU MUST SPECIFY full installation - Must install MinGW-w64           32/64-bit toolchain to compile C, Fortran and C++
  - - You must add both /rtools/bin and something else (forgot) to path 
- run install.packages("Rcpp", type="source")
- run install.packages("RcppArmadillo", type="source")
- run install.packages("mrgsolve_0.5.11.tar.gz", repos=NULL, type="source")
- install dplyr/BH
This could be a major impediment for casual users/noobs

#Part 1 - Vanco IV dosing
##Step 1 - specify a model
Options
- Specify in editor as a .cpp code
- specify as a string 
```{r}
library(mrgsolve)
library(ggplot2)
#load base example model, then strip of data to use as template
cat(readLines("pkpdExample.cpp"), sep="\n")

##modify via editor for a 1cmpt IV - see text changes
cat(readLines("Vanco1cmpt.cpp"), sep="\n")


#Now you must complie using mread()
mod <- mread('Vanco1cmpt')
```

##Step 2 - specify dosing
```{r}

maint <- ev(amt= 1000, ii=12, addl = 1000)
```

##Step 3 - sim/plot
```{r}
bolussim <- mod %>% ev(maint) %>% mrgsim(end = 72) %>% as.data.frame



ggplot(bolussim, aes(x=time, y = CP)) + geom_line()
```


#Part 2 - IV Vanco 2 hour infusion 
##Step 1
mod may be reused.  Only need to respecify the dosing
```{r}

infusion <- ev(amt=1000, rate = 500, ii=12, addl=12)

infusionsim <- mod %>% ev(infusion) %>% mrgsim(end = 72) %>%as.data.frame



ggplot(infusionsim, aes(x=time, y = CP)) + geom_line()
```

#Part 3 - compare dosing regimens
Important - avoid using rbind and iterative solutions to simulating.  rbind requires generating a new df and copying old df's to memory for each iteration. 

Rather than Rbinding, grouping, etc, you can do this by specifying a more complex "event" using expand.ev()

**LOOK expand.ev()** is a very clean function to generate NONMEM type multiple dosing regimens. 
However - this will not work as intended.  it generates an ID with unique dosing for every possible combination of values in vectors. I.E., below code generates 1000 QD and 2000 BID regimens which I do not want to include.
```{r}

multidose <- expand.ev(amt=c(1000,2000), rate=c(500), ii=c(12,24), addl=2)

head(as.data.frame(multidose))
```

Luckily this also allows "adding" of dosing regimens!

```{r}

q12h <- ev(amt= 1000, rate = 500, ID = 1, ii=12, addl = 12)

q24h <- ev(amt=2000, rate = 500, ID = 2, ii=12, addl=12)

both = q12h + q24h



multiregimen <- mod %>% ev(both) %>% mrgsim(end = 72) %>%as.data.frame



ggplot(multiregimen, aes(x=time, y = CP, group = ID)) + geom_line()
```

#Part 4 Population PK 
CV of params: 100 patients, CL 30.13%; V1 22.83%

This requires changing the model completely.  But it is nice to see where the math is coming from, i guess for a n00b.
Added requirements to model - 

```{r}

##We are required to update the model to include $OMEGA and $THETA.  This requires recompilation. 

##OLD
cat(readLines("Vanco1cmpt.cpp"), sep="\n")

##NEW
cat(readLines("Vanco1cmptPOPPK.cpp"), sep="\n")

modPOPPK <- mread('Vanco1cmptPOPPK')


## this takes a covariate n x n matrix as an input for omat

##not clear - order of params in matrix.  going to assume M_ij is the variance for ordered param i = j.  IE. CL is first named param in model so its variance is M(1,1).

##Create variance matrix.  
cov = matrix(c(.3013, 0,0, .2283), 2,2)


popPK <- modPOPPK %>% ev(maint) %>% omat(cov) %>% mrgsim(nid = 100, end = 100) %>% as.data.frame

ggplot(popPK, aes(x=time, y = DV, group = ID)) + geom_line()
```
#Part 5
mrgsolve does not have a covariate class.  Not a big deal, because you may impliment `idata()`, to modify individual parameters.  we can determine these parameters by calculating outside of CPP

AFAIK the covariate class in PKPDsim has no extra functionality besides calling it a covariate class and separating from parameters.  There looks to be more functionality in the pipeline but ron keiser has commented out several functions

```{r}

AGE = 61.1
CRCL = 86.1/70 ##in ml/min/kg
WT = 70 ##kg

ptCL <- (0.67*CRCL + AGE^0.24)*70*0.006

ptcov <- data.frame(CL = ptCL,V = 59)  ### 59 taken from V from model 

covariates <- mod %>% idata_set(ptcov) %>% ev(maint) %>% mrgsim(end = 72) %>% as.data.frame

ggplot(covariates, aes(x=time, y = CP, group = ID, color = factor(ID))) + geom_line()


```

#Extra Fun!
(This is what I originally thought we were going for with covariates)
popPK w/ covariates (assume no covariation between V and CL, I.E. Cov Matrix is a diagnal matrix)
Population average CL/Age in study: 61.1(sd = 16.3) y, CL = 86.1(sd = 55.1) mL/min
```{r}

##generate a normally distributed, random generated patient set
CRCL <- rnorm(10,86.1,55.1)
AGE <- rnorm(10, 61.1, 16.3)
V <- rnorm(10, 59, 13.5)  ### sd extrapolated from CV = 22.83% and mean of 59 from model 
popcov <- data.frame(V, CRCL, AGE, ID = seq(1:10))


###I have to go back to check this equation. 99% sure i screwed something up.  Jimmied to give ballpark correct clearances

popcov <- popcov %>% dplyr::mutate(CL = (0.67*(CRCL/70) + AGE^0.24)*70*0.06)

covariates <- mod %>% idata_set(popcov) %>% ev(maint) %>% mrgsim(end = 72) %>% as.data.frame

ggplot(covariates, aes(x=time, y = CP, group = ID, color = factor(ID))) + geom_line()
```


###Discussion
- mrgsolve is SUBSTANTIALLY faster than PKPDsim.  I simulated 10000 and couldn't get to 5 mississippi. plotting a different story 
- it is equally as easy to build parameter values (based on covariates or popPK) outside of cpp and feed them simultaneously into mrgsim (see #5 and "extra").  This may get slower if we have a big covariate matrix 
