```{r}
library(PKPDsim)
library(ggplot2)
```
Model Plan 
Params(extrapolated): CL = 5.1L/hr; v1 = 59.86 L
1. Dosing regimen (bolus) single patient : 1g q12h
2. Dosing regimen (infusion) single patient : 1 g q12h over 1h
3. Dosing regimen multiple strategies: 1g q12, 0.5g q6h
4. Population PK - CV of params: 100 patients, CL 30.13%; V1 22.83%
5. Correlates: CL(ml/min/kg) = 0.67 * CrCl + AGE^.24; V1 invariant

  Population average CL/Age in study: 61.1(sd = 16.3) y, CL = 86.1(sd   = 55.1) mL/min

###Vanco 1g q12h IV bolus
Potential limitation - new_ode_model() only has 4 models in its default library.  *Why don't we build some models (easy) and push to the repo?*  (FYI - I don't understand GitHub politics or etiquitte)
```{r}
p <- list(CL = 5.1, V  = 60)

pk1 <- new_ode_model("pk_1cmt_iv")

r1 <- new_regimen(amt = 1000,
              times = c(0, 12, 24, 36))

dat <- sim_ode (ode = "pk1",
                par = p,
                regimen = r1)

ggplot(dat, aes(x=t, y=y)) +
  geom_line() +
   facet_wrap(~comp) +
  theme_empty()
```


###Vanco 1g q 12h 2 hour infusion
Notice how easy it is to switch from push to CI - it requires two additional arguements - 28 key strokes - under  new_regimen() 
```{r}
r2 <- new_regimen(amt = 1000,
              times = c(0, 12, 24, 36),
              type = "infusion", t_inf = 2)

dat2 <- sim_ode (ode = "pk1",
                par = p,
                regimen = r2)

ggplot(dat1, aes(x=t, y=y)) +
  geom_line() +
   facet_wrap(~comp) +
  theme_empty()
```

###compare dosing regimens: vanco 1g q12h push vs 1hr infusion 

sim ode() can compute multiple regimens simultaneously, however new.regimen cannot produce multiple regimens at once. However, you can run two different simulations and rbind, create an ID for each simulation, then ggplot(aes(group = ID, simID)).  STILL less hassel than phoenix for small numbers of strategies IMO. 

join_regimen was recently added to address this.  

There is an option to do this via new_regimen_from_nm_dataset - which takes NONMEM datasets and creates regimens. Showcased two chunks below.

Idea to modify - in the spirit of Phoenix.  Allow args to new.regimen to be matrices with columns individual dosing strategies.  Seems like a simple addition, maybe more difficult syntax

```{r}
##must identify bolus vs infusion, will do simply by modifying the ID.  In more complicated scenarios - POP PK, we can just cbind a dosing regimen identifier

dat2$id <- 2
dat3 <- rbind(dat, dat2)

ggplot(dat3, aes(x=t, y=y, group = id)) +
  geom_line() +
   facet_wrap(~comp, scales = 'free') +  ### i freed up the scales - shows that obs is just normalized conc from cmpt1.  
  theme_empty()
```

###Simulating and comparing several simulations simultaneously.

##### The iterativerbind() option is iterative and not gone into detail here.  I believe the time it takes to simulate and view n number of strategies increases by O(n).  Below is an option for using new.regimen.from.nm.dataset

it is slower building a df using Rstudio rather than cell editing in phoenix Nonlinear ME.  For **several** dosing strategies I would suggest building a excel sheet and importing as a df.  I believe PKPDsim is slower for adding simultaneous dose modeling to an existing framework for **several** dosing strategies using new.regimen.from.nm.dataset.  However, running several separate simulations and rbind() may still be faster.
```{r}
fakeNMdf <- data.frame(
    ID = c(1,1,1,1,1,2,2,2,2,2,3,3,3,3,3),
    EVID = c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1),
    Time = c(0,6,12,18,24,0,6,12,18,24,0,6,12,18,24), 
    AMT = c(0,0,0,0,0,500,500,500,500,500, 1000,0, 1000,0, 1000))
      #### I found a bug.   first ID always gets simulated sparsely - that is, output concentration data is only given for 
r4 <- new_regimen_from_nm_dataset(fakeNMdf)

dat4 <- sim_ode (ode = "pk1",
                par = p,
                regimen = r4)

ggplot(dat4, aes(x=t, y=y, group=id)) +
  geom_line() +
   facet_wrap(~comp) +
  theme_empty()
```


###Now we perform population PK
####Population PK - CV of params: 100 patients, CL 30.13%; V1 22.83%
Notice how this is implimented with only 2 added arguements, n_ind and omega(CV matrix)
```{r}
##variation matrix is tricky to make 

omega <- cv_to_omega(par_cv = list(CL = .3013, V = .228), parameters = list(CL = 'CL', V = 'V'))

dat5 <- sim_ode (ode = "pk1",
                par = p,
                n_ind = 100,
                omega = omega,
                regimen = r1)

ggplot(dat5, aes(x=t, y=y, group=id, colour=factor(id))) +
  scale_colour_discrete(guide=FALSE) +
  geom_line() +
  facet_wrap(~comp, scales = 'free') +
  theme_empty()
```

### Model with covariates.

#### Updated - **Meaningful** definition of covariate is redefined based on better understanding the end goal of the project.  We want to feed covariates for a *single clinical patient* **NOT** *pop pk/specify many individual covariates or a pdf for a large pop*

```{r}

##patient params
age  <- new_covariate(value = 80)
CrCl <- new_covariate(value = 80)
wt <- new_covariate(value = 70)
#############################


covmod <- new_ode_model(code = "
                        dAdt[1] = -(((0.67*CrCl + pow(age, 0.24))*0.006*wt)/(V)) * A[1] + rate
                        conc = A[1]/V
                            ",
                     covariates = list("CrCl" = CrCl, "age" = age, 'wt' = wt),
                     cpp_show_code = TRUE,
                     obs = list (cmt = 1, scale = "V"))




dat6 <- sim_ode (ode = "pk4",
                par = p, 
                obs_step_size = .025,  ##### need this, otherwise bolus peaks aren't included
               covariates = list("CrCl" = CrCl, "age" = age, 'wt' = wt),
              'regimen' = r1)


ggplot(dat, aes(x=t, y=y)) +
  scale_colour_discrete(guide=FALSE) +
  geom_line() +
  facet_wrap(~comp, scales = 'free') +
  theme_empty()
                                  
```

## General thoughts - comment?
-Better dosing regimen specification - no clean way of simultaneously simulating several.
-Built in library is very small (4).

## post-hoc thoughts after doing mrgsolve
- mrgsolve allows on-the-fly non- persistent changes in parameters.  Useful for speed in a clinical scenario.  Clinicians want to see results from several different 

## Potential side projects 
- Replicate in Python/Cython
- Push more models/drugs in library (will coincide with our objectives maybe)
- Make shiny app more suited for PharmD PK class
